\normalfont{A bird's-eye view on the habitability of exoplanets via statistical learning techniques}

### 2020-09-22 Manuscript/Manuscript.tex
pub
PUB

\section{Introduction}
The possibility of other form of life on other planets represent an issue that involved scientists as well as philosophers from different millennia of documented human history \cite{papagiannis1985historical}. During the XX century, due to the technological progress, such question moved from a pure speculative approach, as it was previously for Lucretius, Muhammad al-Baqir or Iordanus Brunus, to a more quantitative and scientific method.Furthermore it is interesting to point out that this research was undertaken when the exploration of the Earth was almost completed \cite{fleming2001barrow}. The exploration of space, and the investigation of other planets was largely boosted with the use of telescopes that measure the radiation also outside the visible spectrum and later with the space telescopes such as the NASA's Kepler.
As requested by all form of life in the Earth an habitable planet needs that liquid water can be present \cite{seager2013exoplanet,mckay2014requirements,rothschild2001life}\footnote{Astrobiologist scholars also proposed that other form of life can come up without liquid water (see for instance \cite{rahm2016polymorphism}) however, for simplicity, here only planets that allows liquid water are considered habitable}: this feature that strictly depends by the radiation intensity $I$ is bounded to the power radiation ($P$) produced by the star that host the planet and the respective distance $r$. Indeed by approximating  a star as a point or a perfect sphere, a radiation with a spherical symmetry is produced; thus the integral that gives the power produced $P$:

@@

This relation establish an interval for the distance $r$ where the stellar radiation is not too hot so that the water is the gas phase, and not to low so only the solid state phase is allowed. In the literature this interval is usually called habitable-zone \cite{kasting1993habitable}.

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\textwidth]{Pic/Planets_habitability_Seager.png}
\caption{Seager representation of planet habitability as function of star type (that determinates the stellar activity) and the planet distance from it. Image taken from \cite{seager2013exoplanet}}
\label{Planets_habitability_Seager}
\end{center}
\end{figure}

>>>>

\section{Introduction}
The possibility of other form of life on other planets represent an issue that involved scientists as well as philosophers from different millennia of documented human history \cite{papagiannis1985historical}. During the XX century, due to the technological progress, such question moved from a pure speculative approach, as it was previously for Lucretius, Muhammad al-Baqir or Iordanus Brunus, to a more quantitative and scientific method.Furthermore it is interesting to point out that this research was undertaken when the exploration of the Earth was almost completed \cite{fleming2001barrow}. The exploration of space, and the investigation of other planets was largely boosted with the use of telescopes that measure the radiation also outside the visible spectrum and later with the space telescopes such as the NASA's Kepler.
As requested by all form of life in the Earth an habitable planet needs that liquid water can be present \cite{seager2013exoplanet,mckay2014requirements,rothschild2001life}\footnote{Astrobiologist scholars also proposed that other form of life can come up without liquid water (see for instance \cite{rahm2016polymorphism}) however, for simplicity, here only planets that allows liquid water are considered habitable}. This feature is given, at first glance, by the radiation intensity $I$ that is bounded to the power radiation ($P$) produced by the star that host the planet and the respective distance $r$. Indeed by approximating  a star as a point or a perfect sphere, a radiation with a spherical symmetry is produced; thus the integral that gives the power produced $P$:

@@

Such relation establish an interval for the distance $r$ where the stellar radiation is not too hot so that the water is the gas phase, and not to low so only the solid state phase is allowed. In the literature this interval is usually called habitable-zone \cite{kasting1993habitable}. However this represents a too simplified model since no atmospheric effect is taken in to account: indeed the surface temperature is also influenced by the greenhouse gases that allows to the host radiation to enter inside the planet but limits its escape \cite{seager2013exoplanet,forget2014possible}. Basically this is due to the fact that the greenhouse molecules (CO$_{2}$,vapour, methane...) are excited by the infra-red radiation produced by the planet and remit it in all directions. As consequence the radiation that is emitted by the planet is reduced, and so the temperature is increased. It is worth nothing that in principle the greenhouse is not negative since it allows to the planet to keep part of the heath received from the host star. However this can be a problem if the effect so large that the temperature of the planet abruptly increase: this was probably the fate of Venus that lost his oceans \cite{way2016venus,luger2015extreme,seager2013exoplanet}.

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\textwidth]{Pic/Greenhouse-effect-t2.png}
\caption{The greenhouse effect: the radiation of the host star is captured by the planet and than remitted: this process is limited by the greenhouse molecules of the athomosphere that absorb and re-emit the radiation in all directions. Image taken from \cite{seager2013exoplanet}}
\label{Planets_habitability_Seager}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\textwidth]{Pic/Planets_habitability_Seager.png}
\caption{Seager representation of planet habitability as function of star type (that determinates the stellar activity) and planet distance. Image taken from \cite{seager2013exoplanet}}
\label{Planets_habitability_Seager}
\end{center}
\end{figure}

\clearpage
